---
title: "DATA_605_Assign_12_DThonn"
file: DATA_605_Assign_12_DThonn
name: Daniel Thonn
output:
  html_document: default
  pdf_document: default
course: DATA 605 CUNY MSDA program
---

** DATA_605_Discussion_12_Regression_2 **

http://rpubs.com/danthonn/382369

** Real World Data **



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

``````{r, echo=TRUE}
if (!require(RCurl)) install.packages("RCurl")
if (!require(stringr)) install.packages("stringr")
library(RCurl)
library(stringr)
library(knitr)

if (!require(tidyr)) install.packages("tidyr")
if (!require(dplyr)) install.packages("dplyr")
library(tidyr)
library(dplyr)

if (!require(gvlma)) install.packages("gvlma")
library(gvlma)


```


** Assign_12_Regression_Real_World Data **

The attached who.csv dataset contains real-world data from 2008. The variables included follow.
Country: name of the country
LifeExp: average life expectancy for the country in years
InfantSurvival: proportion of those surviving to one year or more
Under5Survival: proportion of those surviving to five years or more
TBFree: proportion of the population without TB.
PropMD: proportion of the population who are MDs
PropRN: proportion of the population who are RNs
PersExp: mean personal expenditures on healthcare in US dollars at average exchange rate
GovtExp: mean government expenditures per capita on healthcare, US dollars at average exchange rate
TotExp: sum of personal and government expenditures.
1. Provide a scatterplot of LifeExp~TotExp, and run simple linear regression. Do not transform the variables. Provide and interpret the F statistics, R^2, standard error,and p-values only. Discuss whether the assumptions of simple linear regression met.
2. Raise life expectancy to the 4.6 power (i.e., LifeExp^4.6). Raise total expenditures to the 0.06 power (nearly a log transform, TotExp^.06). Plot LifeExp^4.6 as a function of TotExp^.06, and r re-run the simple regression model using the transformed variables. Provide and interpret the F statistics, R^2, standard error, and p-values. Which model is "better?"
3. Using the results from 3, forecast life expectancy when TotExp^.06 =1.5. Then forecast life expectancy when TotExp^.06=2.5.
4. Build the following multiple regression model and interpret the F Statistics, R^2, standard error, and p-values. How good is the model?
LifeExp = b0+b1 x PropMd + b2 x TotExp +b3 x PropMD x TotExp
5. Forecast LifeExp when PropMD=.03 and TotExp = 14. Does this forecast seem realistic? Why or why not?


``````{r, echo=TRUE}

url <- 'https://raw.githubusercontent.com/danthonn/DATA_605_Assign_12_DThonn/master/who.csv'

who1 <- read.csv(url)

str(who1)

glimpse(who1)

head(who1)

kable(head(who1))

```



** Question-1 **

1. Provide a scatterplot of LifeExp~TotExp, and run simple linear regression. Do not transform the variables. Provide and interpret the F statistics, R^2, standard error,and p-values only. Discuss whether the assumptions of simple linear regression met.


``````{r, echo=TRUE}
# Question-1 Answer:


# a simple linear regression - who1
lm_who1 <- lm(who1$LifeExp ~ who1$TotExp)
# a scatter plot - who1
plot(who1$LifeExp, who1$TotExp, main="Scatterplot", 
    xlab="LifeExp", ylab="TotExp", pch=19) 
abline(lm_who1, col = "blue")




# the residuals - who1
hist(resid(lm_who1), main = "Histogram - Residuals", xlab = "residuals")

# summary - who1
summary(lm_who1)



```

Interpretations-1:
Plot: line does not appear linear and is sharply curved upward after LifeExp = 75.
F statistics: large at 65.26.  reject null hypothesis and no relationship between variables.
R^2: R^2 is small at .2577.  Only 25% of response variable explained by the independant variable.
standard error: at 9.3 is large, indicating how much the avg. response variable deviates from the regression line.
p-values: near zero.  probability of observing this relationship is small.


** Question-2 **

2. Raise life expectancy to the 4.6 power (i.e., LifeExp^4.6). Raise total expenditures to the 0.06 power (nearly a log transform, TotExp^.06). Plot LifeExp^4.6 as a function of TotExp^.06, and r re-run the simple regression model using the transformed variables. Provide and interpret the F statistics, R^2, standard error, and p-values. Which model is "better?"

``````{r, echo=TRUE}
# Question-2 Answer:
# a simple linear regression - who2
LifeExp4.6 <- who1$LifeExp^4.6
TotExp0.06 <- who1$TotExp^0.06
lm_who2 <- lm(LifeExp4.6 ~ TotExp0.06)

# scatter plot - who2
plot(TotExp0.06, LifeExp4.6)
abline(lm_who2, col="blue")


# residuals - who2
hist(resid(lm_who2), main = "Histogram - Residuals", xlab = "residuals")

# summary - who2
summary(lm_who2)

```

Interpretations-2:
Plot: regression line does appear linear
F statistics: larger at 507.7 and higher than who1.  
R^2: R^2 is larger at .7283.  73% of response variable explained by the independant variable.
standard error: at 90490000 is large, indicating how much the avg. response variable deviates from the regression line.
p-values:  p-value: < 2.2e-16 is statistically significant.  less than who1 model.
histogram: near normal distribution

Overall: The who2 model is much improved over the who1 model and a good linear regression fit.


** Question-3 **

3. Using the results from 3, forecast life expectancy when TotExp^.06 =1.5. Then forecast life expectancy when TotExp^.06=2.5.


Then forecast life expectancy when TotExp^.06=2.5.
y=-736527910+620060216x
lifeexpectancy=y(1/4.6)



``````{r, echo=TRUE}

life_ex <- function(x)
{   z <- -736527910 + 620060216 * (x)
    z <- z^(1/4.6)
    print(z)
}
# Life expectancy with TotExp^.06 =1.5
life_ex(1.5)
## [1] 63.31153
# Life expectancy with TotExp^.06 =2.5
life_ex(2.5)
## [1] 86.50645



```



** Question-4 **

4. Build the following multiple regression model and interpret the F Statistics, R^2, standard error, and p-values. How good is the model?
LifeExp = b0+b1 x PropMd + b2 x TotExp +b3 x PropMD x TotExp

``````{r, echo=TRUE}
lm_4 <- lm(who1$LifeExp ~ who1$PropMD + who1$TotExp + who1$PropMD*who1$TotExp)
summary(lm_4)

hist(resid(lm_4), main = "Histogram - Residuals", xlab = "residuals")

plot(fitted(lm_4), resid(lm_4))

# Call:
# lm(formula = who1$LifeExp ~ who1$PropMD + who1$TotExp + who1$PropMD * 
#     who1$TotExp)
# 
# Residuals:
#     Min      1Q  Median      3Q     Max 
# -27.320  -4.132   2.098   6.540  13.074 
# 
# Coefficients:
#                           Estimate Std. Error t value Pr(>|t|)    
# (Intercept)              6.277e+01  7.956e-01  78.899  < 2e-16 ***
# who1$PropMD              1.497e+03  2.788e+02   5.371 2.32e-07 ***
# who1$TotExp              7.233e-05  8.982e-06   8.053 9.39e-14 ***
# who1$PropMD:who1$TotExp -6.026e-03  1.472e-03  -4.093 6.35e-05 ***
# ---
# Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
# 
# Residual standard error: 8.765 on 186 degrees of freedom
# Multiple R-squared:  0.3574,	Adjusted R-squared:  0.3471 
# F-statistic: 34.49 on 3 and 186 DF,  p-value: < 2.2e-16

```

Interpretations-4:

1). p-value is less than .05, therefore model is statistically significant. 
2). F-statistic is 34.49 on 3
3). Based on Rsquare only 34.7% of the variability can be explained by the 3 variables.
4). Residuals are right skewed. There, the linear model is not a good fit.

** Question-5 **

5. Forecast LifeExp when PropMD=.03 and TotExp = 14. Does this forecast seem realistic? Why or why not?

LifeExp=(6.277 * 10^1) +(1.497 * 10^3) * PropMD + (7.233 * 10^(-5) *
TotExp - ((6.026 * 10^(-3) * PropMD * TotExp 


``````{r, echo=TRUE}

LifeExp5 <- ( (6.277*10^1) + (1.497*10^3)*.03 + (7.233*10^(-5))* 14 - ((6.026*10^(-3))*0.03*14) ) 
LifeExp5
## [1] 107.6785
 


```

Interpretations-5:
1). The forecast age of 107.6 is an outlier and is unlikely. 
2). The expenditure is also low.


** END **


``````{r, echo=TRUE}

```


